Scan-Centric, Frequency-Based Method for Characterizing Peaks from Direct Injection Fourier transform Mass Spectrometry Experiments

### 2022-04-08 doc/peakcharacterization_manuscript.Rmd
methods around the p-value comparison

::: {custom-style="MDPI_2.2_heading2"}
### Multi-Class F-Statistics
:::

::: {custom-style="MDPI_3.1_text"}
To evaluate the F-statistic between multi-class samples, a much larger set of samples was used.
In this case, 181 matched non-cancer and cancer samples previously used in the HPD detection manuscript [@mitchellNewMethodsIdentify2018].
All 181 samples were peak characterized using the most stringent method below as well as peaks generated from the MSnbase R package [ref], assigned using SMIRFE, and then peaks matched by shared EMFs across samples (see Supplemental materials for a description of peak matching and deciding the most likely EMF).
For each matched peak across samples, we kept the peaks present in at least 25% of samples in both classes of sample (disease, non-disease), and calculated the F-statistic for peaks generated using scan-level peak-characterization and those from MSnbase.
:::

>>>>

::: {custom-style="MDPI_2.2_heading2"}
### Differential Analysis of Large Dataset
:::

```{r load_qcqa}
tar_load(qcqa)
n_out = qcqa$outlier_data %>%
  dplyr::group_by(sample_class) %>%
  dplyr::summarise(n_out = sum(outlier))
n_out_vector = n_out$n_out
names(n_out_vector) = n_out$sample_class
```

::: {custom-style="MDPI_3.1_text"}
To compare p-value changes in a large multi-class sample dataset, we used a much larger set of samples.
We started with the 181 matched non-cancer and cancer samples previously used for HPD detection and lipidomics of non-small-cell-lung-cancer (NSCLC) [@mitchellNewMethodsIdentify2018; @mitchellUntargetedLipidomicsNonSmall2021].
All samples were characterized using the full scan-centric workflow, as well as peaks generated from the MSnbase R package [@gattoMSnbaseBioconductorPackage2012; @gattoMSnbaseEfficientElegant2020], assigned using SMIRFE, and then peaks matched by shared EMFs across samples (see other methods for a description of peak matching and deciding the most likely EMF).
After extraction of the scan-centric IMF peak locations across all samples, Xcalibur and MSnbase peaks were matched within each sample, and locations and intensities extracted.

Of the starting 182 samples, five did not finish during scan-centric characterization due to not having any scans left during normalization.
In addition, some of the samples were acquired multiple times, creating duplicate samples.
Not knowing which sample run was most appropriate to use, we removed all of the duplicated samples.
Using information-content-informed Kendall-tau correlation, we compared the median correlations of each sample to all others in the same disease class for outliers, and removed `r n_out_vector["cancer"]` cancer and `r n_out_vector["non-cancer"]` samples.

Each samples intensities were normalized by dividing by the median intensities from that sample.
Only those peaks present in 50% of both the non-cancer and cancer samples were kept for differential analysis.
Differential analysis used the logged intensity values, with p-values calculated using rowttests function from the genefilter Bioconductor package (v1.76.0) [@gentlemanGenefilterMethodsFiltering2021], removing any missing values before calculation.
P-values were adjusted using the Benjamini-Hochberg method in the p.adjust function from the base R stats package (v4.1.0) [@rcoreteamLanguageEnvironmentStatistical2021a].

For comparisons across peak sources, we converted both the raw p-values to log-p-values by calculating -1 * Log10(p-value).
The reference peak source is the raw scan-centric p-values.
:::
